### 2015-06-12 ApresentacoesArtigos/quali/quali.tex
delineada geral na monografia de qualificacao

%\chapter{Materials}\label{sec:mat}
\chapter{Materiais}\label{sec:mat}
%\section{The Gmane public database of email lists (benchmarks)}
\section{O banco de dados públicos sobre listas de email (benchmark)}
\section{Facebook, Twitter, Participa.br, Cidade Democrática, AA}
%\section{My own social networks?}
\section{Minhas próprias redes sociais}
Considerações sobre o uso de minhas anotações para
pesquisa sobre mim e meu mundo.

\chapter{Methods}
\section{Circular statistics}
\section{Erdös Sectioning}
\section{PCA of measures along time}
\section{Kolmogorov-Smirnoff test for texts produced by sectors}
\section{Audiovisualization of data}
\section{Typological considerations}
\section{Semantic web}
\subsection{OWL ontology construction}

\chapter{Results}
\section{Time stability in human interaction networks}
\section{Semantic web}
\subsection{Linked data}
\subsection{RDF data conversion of data into linked data}
\subsection{Published linked data and OWL ontologies}
\section{Harnessing}
\subsection{Social percolation procedures}
\subsection{Recommendation systems for the enrichment of semantic navigation}
\subsection{Understanding the social being}
Scale free as the consequence of $T^2$ signal, fractal, constant, with three primitive parts and greater specialization. Gradus ad Parnassum

\chapter{Finished and planed tasks, chronogram}
% colocar comparativos aqui
\section{Documents}
\subsection{To be finished}
\subsubsection{Anthropological physics}
The study of human systems raises conceptual
and ethical issues that require anthropological considerations.
There are two immediate routes to this concepts:
\begin{itemize}
	\item What data should or can be used?
	\item Can one experiment in a network of humans? In which context?
\end{itemize}

@@

\chapter{Conclusions}

>>>>

%\chapter{Materials}\label{sec:mat}
\chapter{Materiais}\label{sec:mat}
%\section{The Gmane public database of email lists (benchmarks)}
\section{O banco Gmane de dados públicos sobre listas de email (benchmark)}

Mensagens de lista de email foram obtidas do arquivo Gmane~\cite{gmanePack},
que consiste em mais de 20 mil listas de email e mais de 130 milhões de mensagens~\cite{GMANEwikipedia}.
Estas listas cobrem uma variedade de assuntos, em especial
relacionados à tecnologia. O arquivo pode ser descrito como
um corpus com metadados de emails, que incluem hora e lugar
de envio, nome e email do remetente. O uso do GMANE para
pesquisa científica é incidente no estudo de listas isoladas
e de inovações lexicais~\cite{GMANE2,bird}.

\section{Facebook, Twitter, Participa.br, Cidade Democrática, AA}
Embora as redes de email tenham sido usadas como referência
na observação de propriedades gerais, outras fontes
foram analisadas:

\begin{itemize}
	\item Redes do Facebook: 8 são usadas como referência em~\cite{timeS}, mas dezenas, talvez algumas centenas, foram observadas nos experimentos da Seção~\ref{sec:exp}.
	\item milhares de tweets (talvez alguns milhões), dados de portais
 	\item mecanismos participativos como o Participa.br, Cidade Democrática e o AA.
\end{itemize}

\section{Minhas próprias redes sociais}
Considerações sobre o uso de minhas anotações para
pesquisa sobre mim e meu mundo.

\chapter{Métodos}
\section{Estatística circular}
Messages were counted along time with respect to seconds, minutes, hours, days of the week, days of the month, and months of the year. This resulted in histograms from which patterns could be drawn. The ratio $\frac{b_h}{b_l}$ between the highest and lowest incidences on the histograms served as a hint of how close the observed distribution is to a uniform distribution.

The average and the dispersion were taken using circular statistics, in which each $measurement$ (data point) is represented as a complex number with modulus equal to one, $z=e^{i\theta}=\cos(\theta)+i\sin(\theta)$, where $\theta=measurement\frac{2\pi}{period}$. The moments $m_n$, lengths of moments $R_n$, mean angle $\theta_\mu$, and rescaled mean angle $\theta_\mu'$ are defined as:

\begin{align}\label{eq:cmom}
	m_n&=\frac{1}{N}\sum_{i=1}^N z_i^n \nonumber\\
	R_n&=|m_n|\\
	\theta_\mu&=Arg(m_1) \nonumber \\
	\theta_\mu'&=\frac{period}{2\pi} \theta_\mu \nonumber
\end{align}

$\theta_\mu'$ is used as the measure of location. Dispersion is measured using the circular variance $Var(z)$, the circular standard deviation $S(z)$, and the circular dispersion $\delta(z)$:

\begin{align}\label{eq:cmd}
	Var(z)&=1 - R_1 \nonumber\\
	S(z)&= \sqrt{-2\ln(R_1)}\\
	\delta(z)&=\frac{1-R_2}{2 R_1^2} \nonumber
\end{align}

As expected, a positive correlation was found in all $Var(z)$, $S(z)$ and $\delta(z)$ dispersion measures, as can be noticed in Section~\ref{si:circ} of the Supporting Information, and $\delta(z)$ was preferred in the discussion of results.

\subsection{Formação das redes de interação}\label{sec:intNet}
Interaction networks can be modeled both weighted or unweighted, both directed or undirected~\cite{bird,newmanCommunityDirected,newmanCommunity2013}.
Networks in this article are directed and weighted, the more informative of trivial possibilities, i.e. we did not investigate directed unweighted, undirected weighted, and undirected unweighted representations of the interaction networks. 
The networks were obtained as follows: a direct response from participant B to a message from participant A yields an edge from A to B, as information went from A to B. The reasoning is: if B wrote a response to a message from A, he/she read what A wrote and formulated a response, so B assimilated information from A, thus $A \rightarrow B$. Inverting edge direction yields the status network: B read the message and considered what A wrote worth responding, giving status to A, thus $B\rightarrow A$. This article uses the information network as described above and depicted in Figure~\ref{formationNetwork}. Edges in both directions are allowed. Each time an interaction occurs, one is added to the edge weight. Selfloops were regarded as non-informative and discarded. These human social interaction networks are reported in the literature as exhibiting scale-free and small world properties, as expected for (some) social networks~\cite{bird,newmanBook}.

\begin{figure}[!h]
	\centering
	\includegraphics[width=0.5\textwidth]{figs/criaRede__}
	\caption{Formation of interaction network from email messages. Each vertex represents a participant. A reply message from participant B to a message from participant A is regarded as evidence that B received information from A and yields a directed edge. Multiple messages add ``weight'' to a directed edge. Further details are given in Section~\ref{intNet}.}
	\label{formationNetwork}
\end{figure}

Edges can be created from all antecedent message authors on the message-response thread to each message author.
We only linked the immediate antecedent to the new message author, both for simplicity and for the valid objection that in adding two edges, $x\rightarrow y$ and $y\rightarrow z$, there is also a weaker connection between $x$ and $z$. Potential interpretations for this weaker connection are: double length, half weight or with one more ``obstacles''. This suggests the adequacy of centrality measurements to account for the connectivity with all nodes, such as betweenness centrality, eigenvector centrality and accessibility~\cite{luMeasures,access}.

\section{Seccionamento de Erdös}
In scale-free networks, the peripheral, intermediary and hubs sectors can be derived from a comparison with an Erd\"os-R\'enyi network with the same number of edges and vertices~\cite{3setores}, as depicted in Figure~\ref{fig:setores}. We shall refer to this procedure as \emph{Erd\"os sectioning}, with the resulting sectors being referred to as \emph{Erd\"os sectors} or \emph{primitive sectors}.

The degree distribution $\widetilde{P}(k)$ of an ideal
scale-free network $\mathcal{N}_f$ with $N$ vertices and $z$ edges has less
average degree nodes than the distribution $P(k)$ of an Erd\"os-R\'enyi
network with the same number of vertices and edges. Indeed, we define in this work the intermediary sector of a network to be the set of all the nodes whose degree is less abundant in the real network than on the Erd\"os-R\'enyi model:

\begin{equation}\label{criterio}
	\widetilde{P}(k)<P(k) \Rightarrow \text{k is intermediary degree}
\end{equation}

If $\mathcal{N}_f$ is directed and has no self-loops, the probability
of an edge between two arbitrary vertices is $p_e=\frac{z}{N(N-1)}$.
A vertex in the ideal Erd\"os-R\'enyi digraph with the same number of vertices and edges, and thus the same probability $p_e$ for the presence of an edge, will have degree $k$ with probability:

\begin{equation}
	P(k)=\binom{2(N-1)}{k}p_e^k(1-p_e)^{2(N-1)-k}
\end{equation}

The lower degree fat tail represents the border vertices, i.e. the peripheral sector or periphery where $\widetilde{P}(k)>P(k)$ and $k$ is lower than any intermediary sector value of $k$. The higher degree fat tail is the hub sector, i.e. $\widetilde{P}(k)>P(k)$ and $k$ is higher than any intermediary sector value of $k$. The reasoning for this classification is: 1) vertices so connected that they are virtually inexistent in networks connected at pure chance (e.g. without preferential attachment) are correctly associated to the hubs sector. Vertices with very few connections, which are way more abundant than expected by pure chance, are assigned to the periphery. Vertices with degree values predicted as the most abundant if connections are created by pure chance, near the average, and less frequent in the real network, are classified as intermediary.

\begin{figure}[!h]
	\centering
	\includegraphics[width=0.5\textwidth]{figs/fser_}
	\caption{Degree distribution of scale-free and Erd\"os-R\'enyi ideal networks.
		The latter has more intermediary vertices, while the former has more peripheral and hub vertices.
		The sector borders are defined by the two intersections $k_L$ and $k_R$ of the connectivity distributions.
		Characteristic degrees are in the compact intervals: $[0,k_L]$, $(k_L,k_R]$, $(k_R,k_{max}]$ for the Erd\"os sectors (periphery, intermediary and hubs).}
		\label{fig:setores}
\end{figure}
i	To ensure statistical validity of the histograms, bins can be chosen to contain at least $\eta$ vertices of the real network. Thus, each bin, starting at degree $k_i$, spans $\Delta_i=[k_{i},k_{j}]$ degree values, where $j$ is the smallest integer with which there are at least $\eta$ vertices with degree larger than or equal $k_i$, and less than or equal $k_{j}$. This changes equation~\ref{criterio} to:

\begin{equation}\label{criterio2}
	\sum_{x=k_i}^{k_j} \widetilde{P}(x) < \sum_{x=k_i}^{k_j} P(x) \Rightarrow \text{i is intermediary}
\end{equation}

If strength $s$ is used for comparison, $P$ remains the same, but $P(\kappa_i)$ with $\kappa_i=\frac{s_i}{\overline{w}}$ should be used for comparison, with $\overline{w}=2\frac{z}{\sum_is_i}$ the average weight of an edge and $s_i$ the strength of vertex $i$. For in and out degrees ($k^{in}$, $k^{out}$) comparison of the real network should be made with:
\begin{equation}
	\hat{P}(k^{way})=\binom{N-1}{k^{way}}p_e^k(1-p_e)^{N-1-k^{way}}
\end{equation}

\noindent where \emph{way} can be \emph{in} or \emph{out}. In and out strengths ($s^{in}$, $s^{out}$) are divided by $\overline{w}$ and compared also using $\hat{P}$. Note that $p_e$ remains the same, as each edge yields an incoming (or outgoing) edge, and there are at most $N(N-1)$ incoming (or outgoing) edges, thus $p_e=\frac{z}{N(N-1)}$ as with the total degree.

In other words, let $\gamma$ and $\phi$ be integers in the intervals $1 \leq \gamma \leq 6$, $1 \leq \phi \leq 3$, and each of the basic six Erd\"os sectioning possibilities $\{E_{\gamma}\}$ have three Erd\"os sectors $E_{\gamma}= \{e_{\gamma, \phi} \}$ defined as:

\begin{alignat}{3}\label{eq:part}
	e_{\gamma,1}&=\{\;i\;|\;\overline{k}_{\gamma,L}\geq&&\overline{k}_{\gamma,i}\} \nonumber \\
	e_{\gamma,2}&=\{\;i\;|\;\overline{k}_{\gamma,L}<\;&&\overline{k}_{\gamma,i}\leq\overline{k}_{\gamma,R}\} \\ 
	e_{\gamma,3}&=\{\;i\;|\;&&\overline{k}_{\gamma,i}<\overline{k}_{\gamma,R}\} \nonumber
\end{alignat}

\noindent where $\{\overline{k}_{\gamma,i}\}$ is:

\begin{equation}
	\begin{split}
		\overline{k}_{1,i}&=k_i \\
		\overline{k}_{2,i}&=k_i^{in} \\
		\overline{k}_{3,i}&=k_i^{out} \\
		\overline{k}_{4,i}&=\frac{s_i}{\overline{w}} \\
		\overline{k}_{5,i}&=\frac{s_i^{in}}{\overline{w}} \\
		\overline{k}_{6,i}&=\frac{s_i^{out}}{\overline{w}} \\
	\end{split}
\end{equation}

\noindent and both $\overline{k}_{\gamma,L}$ and $\overline{k}_{\gamma,R}$ are found using $P(\overline{k})$ or $\hat{P}(\overline{k})$ as described above.

Since different metrics can be used to identify the three types of vertices, compound criteria can be defined. For example, a very stringent criterion can be used, according to which a vertex is only regarded as pertaining to a sector if it is so for all the metrics. After a careful consideration of possible combinations, these were reduced to six:

\begin{itemize}
	\item Exclusivist criterion $C_1$:  vertices are only classified if the class is the same according to all metrics. In this case, vertices classified (usually) do not reach 100\%, which is indicated by a black line in Figure~\ref{fig:sectIL}.
	\item Inclusivist criterion $C_2$: a vertex has the class given by any of the metrics. Therefore, a vertex may belong to more than one class, and the total number of members may exceed 100\%, which is indicated by a black line in Figure~\ref{fig:sectIL}.
	\item Exclusivist cascade $C_3$: vertices are only classified as hubs if they are hubs according to all metrics. Intermediary are the vertices classified either as intermediary or hubs with respect to all metrics. The remaining vertices are regarded as peripheral.
	\item Inclusivist cascade $C_4$: vertices are hubs if they are classified as so according to any of the metrics. The remaining vertices are classified as intermediary if they belong to this category for any of the metrics. Peripheral vertices will then be those which were not classified as hub or intermediary with any of the metrics. 
	\item Exclusivist externals $C_5$: vertices are only hubs if they are classified as such according to all the metrics. The remaining vertices are classified as peripheral if they fall into the periphery or hub classes by any metric. The rest of the nodes are classified as intermediary.
	\item Inclusivist externals $C_6$: hubs are vertices classified as hubs according to any metric. The remaining vertices will be peripheral if they are classified as such according to any metric. The rest of the vertices will be intermediary vertices.
\end{itemize}

Using equations~\ref{eq:part}, these compound criteria $C_\delta$, with $\delta$ integer in the interval $1<\delta<6$ can be described as:

%\begin{alignat}{3}
\begin{equation}
	\begin{split}
		%\begin{multline}
		C_1&=\{c_{1,\phi}=\left\{i\mid i\;\in e_{\gamma,\phi}, \;\forall\; \gamma\}\right\} \\
		C_2&=\{c_{2,\phi}=\left\{i\mid \exists \;\;\gamma: i \in e_{\gamma,\phi}\}\right\} \\
		C_3&=\{c_{3,\phi}=\left\{i\mid i\;\in e_{\gamma,\phi'}, \;\forall\; \gamma,\;\forall\;\phi'\geq \phi\}\right\} \\
		C_4&=\{c_{4,\phi}=\left\{i\mid i\;\in e_{\gamma,\phi'}, \;\forall\; \gamma,\;\forall\;\phi'\leq \phi\}\right\} \\
		C_5&=\{c_{5,\phi}=\left\{i\mid i\;\in e_{\gamma,\phi'}, \;\forall\; \gamma,\right.\\
																																	&\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \left.\;\forall\;(\phi'+1)\%4\leq (\phi+1)\%4\}\right\} \\
		C_6&=\{c_{6,\phi}=\left\{i\mid i\;\in e_{\gamma,\phi'}, \;\forall\; \gamma,\right.\\
																																	&\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \left.\;\forall\;(\phi'+1)\%4\geq (\phi+1)\%4\}\right\} \\
		%\end{multline}
	\end{split}
\end{equation}
%\end{alignat}

The simplification of all the compound possibilities to the small set listed above can be formalized in strict mathematical terms, but this was considered out of the scope for current interests. It is worth noting that the exclusivist cascade is the same sectioning of an inclusivist cascade from periphery to hubs, but with inverted order of sectors precedence. These compound criteria can be used to examine network sections considering all degrees and strengths and are specially useful in the case of a low number of messages, such as in Section~\ref{si:frac} of the Supporting Information.

\section{Média e desvio do PCA ao longo do tempo}
A Análise de Componentes Principais (PCA é a sigla consagrada, do inglês Principal Component Analysis) foi usada para observar a estabilidade
na formação das componentes. Mais especificamente: 1) quais as medidas que contribuiem para cada componente e em que medida;
2) quanto da dispersão está concentrada em cada componente.

\subsection{Medidas consideradas e acrescentadas}
The topology of the networks was studied using Principal Component Analysis (PCA~\cite{pca}) with a small selection of the most basic and fundamental measurements for each vertex, as follows:

\begin{itemize}
	\item Degree     $k_i$: number of edges linked to vertex $i$.
	\item In-degree  $k_i^{in}$: number of edges ending at vertex $i$.
	\item Out-degree $k_i^{out}$: number of edges departing from vertex $i$.
	\item Strength $s$: sum of weights of all edges linked to vertex $i$.
	\item In-strength $s_i^{in}$: sum of weights of all edges ending at vertex $i$.
	\item Out-strength $s_i^{out}$: sum of weights of all edges departing from vertex $i$.
	\item Clustering coefficient $cc_i$: fraction of pairs of neighbors of $i$ that are linked.  The standard clustering coefficient for undirected graphs was used.
	\item Betweenness centrality $bt_i$: fraction of geodesics that contain vertex $i$. The betweenness centrality index considered directions and weight, as specified in~\cite{faster}.
\end{itemize}

In order to capture symmetries in the activity of participants, the following metrics were introduced for a vertex $i$:

\begin{itemize}
	\item Asymmetry: $asy_i=\frac{k_i^{in}-k_i^{out}}{k_i}$.
	\item Mean of asymmetry of edges: $\mu_i^{asy}=\frac{\sum_{j\in J_i} e_{ji}-e_{ij}}{|J_i|=k_i}$, where $e_{xy}$ is 1 if there is an edge from $x$ to $y$, and $0$ otherwise. $J_i$ is the set of neighbors of vertex $i$, and $|J_i|=k_i$ is the number of neighbors of vertex $i$.
	\item Standard deviation of asymmetry of edges: $\sigma_i^{asy}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{asy} -(e_{ji}-e_{ij}) ]^2  }{k_i}  }$.
	\item Disequilibrium: $dis_i=\frac{s_i^{in}-s_i^{out}}{s_i}$.
	\item Mean of disequilibrium of edges: $\mu_i^{dis}=\frac{\sum_{j \in J_i}\frac{w_{ji}-w_{ij}}{s_i}}{k_i}$, where $w_{xy}$ is the weight of edge $x\rightarrow y$ and zero if there is no such edge.
	\item Standard deviation of disequilibrium of edges: $\sigma_i^{dis}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{dis}-\frac{(w_{ji}-w_{ij})}{s_i}]^2}{k_i}}$.
\end{itemize}

\section{Teste de Kolmogorov-Smirnoff para os textos produzidos por cada setor}

Be $F_{1,n}$ and $F_{2,n'}$ two empirical distribution functions, where $n$ and $n'$ are the number of observations on each sample. The two-sample Kolmogorov-Smirnov test rejects the null hypothesis if:
\begin{equation}\label{eq:ks}
D_{n,n'} > c(\alpha)\sqrt{\frac{n+n'}{nn'}}
\end{equation}

where $D_{n,n'}=sup_x[F_{1,n}-F_{2,n'}]$ and $c(\alpha)$ is given for each level of $\alpha$:

\begin{table}[H]
\centering
\begin{tabular}{|l||c|c|c|c|c|c|}\hline
$\alpha$    & 0.1  & 0.05 & 0.025 & 0.01 & 0.005 & 0.001 \\\hline
$c(\alpha)$ & 1.22 & 1.36 & 1.48  & 1.63 & 1.73  & 1.95  \\\hline
\end{tabular}
\end{table}

\subsection{Adaptação}
We need to compare empirical distribution functions, so $D_{n,n'}$ is given, as are $n$ and $n'$. Therefore, as all terms in equation~\ref{eq:ks} are positive and $c(\alpha)$ can be isolated:

\begin{equation}\label{eq:ks2}
c(\alpha) < \frac{D_{n,n'}}{\sqrt{\frac{n+n'}{nn'}}} = c'(\alpha)
\end{equation}

Tables~\ref{tab:kolSub}-\ref{tab:kolPctInter} are populated with values for $c'(\alpha)$. When $c'(\alpha)$ is high, low values of $\alpha$ are possible for the test to reject the null hypothesis. Therefore, when $c'(\alpha)$ is greater than $\approx 1.7$, it is reasonable to assume that $F_{1,n}$ and $F_{2,n'}$ differ.

%\section{Audiovisualization of data}
\section{Audiovisualização de dados}

Networks were visualized with animations, image galleries and online gadgets developed specifically for this research~\cite{animacoes,galGMANE,appGMANE}. Such visualizations were crucial to guide research into the most important features of network evolution. Furthermore, the size of the three Erd\"os sectors could be visualized in a timeline fashion. Visualization of network structure was especially useful in the inspection of data and derived structures from the email lists.

%\section{Typological considerations}
\section{Considerações tipológicas}
Por enquanto, isso para nós quer dizer:
\begin{itemize}
	\item Consideração sobre o potencial estigmatizante da classificação.
	\item Considerações sobre o indivíduo ou a rede que é classificada, i.e. recebe um tipo.
\end{itemize}

\section{Web semântica}
Para a formalização de conceitualizações, e
para formatos de dados apropriados para armazenamento,
compartilhamento e referência,
foram adotadas as recomendações de linked data
da W3C~\cite{rdf,owl,rdfs}.

\subsection{A construção de ontologias OWL}
Foram construídas ontologias e vocabulários a partir de entrevistas (OBS e VBS principalmente). Um método de levantamento de ontologia orientado aos dados surgiu, potencialmente útil a todos os portais e software em necessidade de ontologias, e foi responsável por 2 ontologias (OPA dados e OCD).

\subsection{A triplificação de dados relacionais}
Para disponibilização e uso de dados de diferentes fontes, foram
feitos pequenos programas de computador (scripts) para
acessar dados relacionais e escrever triplas RDF.

\chapter{Resultados}
%\section{Time stability in human interaction networks}
\section{Estabilidade temporal em redes de interação humana}
Explicitados em~\cite{timeS}.
\section{Criação da nuvem brasileira de dados participativos}
%\subsection{RDF data conversion of data into linked data}
Primeiras ontologias, aplicações, dados, publicações, etc.
Produtos 3,4,5 e artigo sobre OPS.
\subsection{Conversão de dados relacionais para RDF}
Roteiros de triplificação do Participa.br, Cidade Democrática, AA.
%\subsection{Published linked data and OWL ontologies}
\subsection{Síntese de ontologias OWL}
Scripts para síntese, seja livre ou a partir dos dados.
\subsection{Ontologias e dados publicados}
datahub.io, MP, arXiv.

\section{Aproveitamento}
\subsection{Procedimentos de percolação social}
%\subsection{Recommendation systems for the enrichment of semantic navigation}
Procedimentos cíclicos e procedimentos efèmeros.
Experimentos paradigmáticos:
\begin{itemize}
	\item Crowdfunding para a pesquisa de redes (Vaca do fim do mundo).
	\item Escrita para os de maior betweenness e de maior closeness, mostrar que não bate para as listas.
\end{itemize}

\subsection{Sistemas de recomendação para o enriquecimento da navegação semântica de recursos}
%\subsection{Understanding the social being}
Como descrito no produto 4 PNUD. Repassar caracterização básica e direcionar ao escrito.

\subsection{Compreensão sobre a entidade social}
Scale free as the consequence of $T^2$ signal, fractal, constant, with three primitive parts and greater specialization. Gradus ad Parnassum.

%\chapter{Finished and planed tasks, chronogram}
\chapter{Afazeres finalizados, planejados e prontos; cronograma}
% Colocar tabela com os comparativos
% Referenciar para complementos no texto.
\section{Experimentos}\label{sec:exp}
\section{Documents}
\subsection{To be finished}
\subsubsection{Anthropological physics}
The study of human systems raises conceptual
and ethical issues that require anthropological considerations.
There are two immediate routes to this concepts:
\begin{itemize}
	\item What data should or can be used?
	\item Can one experiment in a network of humans? In which context?
\end{itemize}

@@

\chapter{Conclusões}
Há, a princípio, uma confirmação de que os conhecimentos de redes complexas
possuem aplicações diversas e potencialmente benéficas para o participante.
Por exemplo, os experimentos apresentaram modificações da estrutura social para comportar a pesquisa, e podem ser usados para comportar outros empreendimentos. Os estudos de estabilidade e diferenciação em redes de interação humana apontam na direção de tipologias de redes e de participantes, com base nos setores de Erdös.
Um legado de dados ligados e abertos é conveniente para apresentar estes dados às comunidades acadêmicas e interessadas nas aplicações, para o qual foram adiantadas ontologias, vocabulários, rotinas de conversão de dados relacionais em RDF e os dados em si.

Uma direção simples
é focar no \emph{Complex Networks Gradus ad Parnassum},
que é um manual para introdução à àrea através do benefício 
das redes complexas para o indivíduo pela ação dele próprio.
Uma via menos pedagógica, mas mais fundamental, é
explorar as estabilidades encontradas: até que número de
agentes a distribuição dos setores e a formação do PCA se mantém?;
como caracterizar a intermitência dos agentes enquanto a distribuição
grau é estável?

Há, entretanto, considerações teóricas profundas sobre a área,
com implicações sobre a própria constituição das redes complexas.
Ao mesmo tempo, os métodos utilizados também já parecem novos,
e está faltando uma capacidade de apresentar isso para as revistas.
Talvez por falta de equipe adequada para realizar o trabalho,
talvez por falta de prática do doutorando na escrita e encaminhamento
acadêmico.
